TacticToe: Learning to Prove with Tactics

### 2018-08-21 tactictoe_jv.tex
proof plan citations

\textsf{SEPIA}~\cite{DBLP:conf/cade/GransdenWR15} can
generate
proof scripts from previous \coq proof examples.
Its strength lies in the ability to produce likely sequences
of tactics for solving domain specific goals. It operates by creating a model
for common sequences of tactics for a specific library.
This means that in order to propose the next tactic, only the previously
called tactics
are considered.
%intuition tells us that it should not be the main factor
Our algorithm, on the other hand, relies mainly on the characteristics of the
current goal
to decide
which tactics to apply next. In this way, our learning mechanism has to
rediscover why each
tactic was applied for the current subgoals. It may lack some useful bias for
common sequences
of tactics, but is more reactive to subtle changes. Indeed, it can be trained
on a large library and only tactics relevant to the current subgoal will be
selected.
Concerning the proof search, \textsf{SEPIA}'s %simple
breadth-first search is replaced by MCTS which allows for supervised learning
guidance in the exploration of the search tree.
Finally, \textsf{SEPIA} was evaluated on three chosen parts of the
\coq library demonstrating that it globally outperforms individual \coq
tactics. Here, we demonstrate the competitiveness of our system against
\eprover on the \holfour standard library.

@@

A closely related approach is proof planning that was firstly describe in 
~\cite{proofplan}. ~\cite{oyster}
The philosophy of proof planning is to create (automatically or not) a few 
tactics can prove a wide range of theorems. In this respect, if you consider 
MCTS to be a proof plan then we have produced one general tactic that learns 
from examples and has a large success rate.
But proof planning is also concerned with tactic synthesis where new proof 
plans (tactics in \holfour) may be discovered automatically from examples.
From this point of view, \tactictoe does not look at proof patterns to infer 
new tactics and creates new tactics only by abstracting arguments. Therefore, 
it would be interesting to see how including tactic synthesis or even 
human-designed proof plans (such as rippling~\cite{rippling}) in the database 
of tactics $\mathbb{D}_{\mathit{tactic}}$.
Proof planning is a meta approach to tactics constructing more and 
more general tactics by sharing common proof patterns between different tactics.
Comparatively, our approach tries to automate the construction of proof plans 
by letting MCTS policy decide which tactics is the best in each situation. Our 
approach can be thought as one general proof plan only guided by previous human 
proofs. 
From another angle, if we think of proof plans as clever more general specific 
tactics (like inductionless induction\cite{})
our approach does not try to synthetize proof plan automatically but just 
re-use them if they already exists.

%However, we simply create a single label for each tactic call whereas each of
%its
%arguments is treated independently in \textsf{ML4PG}.
%Our choice is motivated by the k-NN algorithm already used in
%\holyhammer for the a of theorems.
%, which can be easly adapted to the selection of tactics but would be improved
%by a
%version that can learn to predict tactics and their arguments independently.

%A recent improvement on this system~\cite{}\todo{Cezary: could not remember
%the
%reference that you gave me here i think it was CICM 2017} includes the
%possibility to
%substitute theorems appearing in tactics by ones belonging to the same
%cluster.
%In this work, we are also able to predict theorems appearing as arguments of
%tactics but we rely only on information from the current goal. Combining the
%two approaches by predicting from the goal and from the original tactic
%arguments is more difficult but is likely to produce more suitable lemmas.
%
Proof patching \cite{RingerYLG18} is an approach that attempts to learn the
proof changes necessary for a modification of a prover library or system. The
approach is hard to apply for new proofs. Following this idea, \tactictoe
could also try to learn from different versions of the prover library.

>>>>

\textsf{SEPIA}~\cite{DBLP:conf/cade/GransdenWR15} can
generate
proof scripts from previous \coq proof examples.
Its strength lies in the ability to produce likely sequences
of tactics for solving domain specific goals. It operates by creating a model
for common sequences of tactics for a specific library.
This means that in order to propose the next tactic, only the previously
called tactics
are considered.
Our algorithm, on the other hand, relies mainly on the characteristics of the
current goal
to decide
which tactics to apply next. In this way, our learning mechanism has to
rediscover why each
tactic was applied for the current subgoals. It may lack some useful bias for
common sequences
of tactics, but is more reactive to subtle changes. Indeed, it can be trained
on a large library and only tactics relevant to the current subgoal will be
selected.
Concerning the proof search, \textsf{SEPIA}'s %simple
breadth-first search is replaced by MCTS which allows for supervised learning
guidance in the exploration of the search tree.
Finally, \textsf{SEPIA} was evaluated on three chosen parts of the
\coq library demonstrating that it globally outperforms individual \coq
tactics. Here, we demonstrate the competitiveness of our system against
\eprover on the \holfour standard library.

@@

A closely related approach is proof planning that was firstly describe in 
~\cite{DBLP:conf/cade/Bundy88}. 
Proof planners have been implemented to facilitate interactive theorem proving 
development in Oyster~\cite{DBLP:conf/cade/BundyHHS90},  
Omega~\cite{DBLP:conf/cade/BenzmullerCFFHKKKMMSSS97} and 
\isabelle~\cite{DBLP:conf/cade/DixonF03}.
The philosophy of proof planning is to create (automatically or not) a few 
tactics can prove a wide range of theorems. In this respect, if you consider 
MCTS to be a proof plan then we have produced one general tactic that learns 
from examples and has a large success rate.
But proof planning is also concerned with tactic synthesis where new proof 
plans (tactics in \holfour) may be discovered automatically from examples.
From this point of view, \tactictoe does not look at proof patterns to infer 
new tactics and creates new tactics only by abstracting arguments. Therefore, 
it would be interesting to see how including tactic synthesis or even 
human-designed proof plans in the database 
of tactics $\mathbb{D}_{\mathit{tactic}}$.
Proof planning is a meta approach to tactics constructing more and 
more general tactics by sharing common proof patterns between different tactics.
Comparatively, our approach tries to automate the construction of proof plans 
by letting MCTS policy decide which tactics is the best in each situation. Our 
approach can be thought as one general proof plan only guided by previous human 
proofs. 
our approach does not try to synthetize proof plan automatically but just 
re-use them if they already exists.

Proof patching \cite{RingerYLG18} is an approach that attempts to learn the
proof changes necessary for a modification of a prover library or system. The
approach is hard to apply for new proofs. Following this idea, \tactictoe
could also try to learn from different versions of the prover library.
